---
title: "Building a linear model"
author: "Fredrik Öhlander"
date: "2024-01-28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

In this project the packages 'ggplot2' 'tibble' was used. The code for
installing is provided below:

```{r}
install.packages("ggplot2")
library(ggplot2)
install.packages("tibble")
library(tibble)
```


#The function contains 4 inputs. 'data', 'dep', 'indep', and 'intercept'.
#'data' is the dataset you wish to use to construct your model.
#'dep' is the name for the vector corresonding to the dependent variable.
#'indep' is a vector of names for the vectors corresponding to the independent
#variables.
#'intercept' determines whether the model has an intercept or not.
#Its output is a list consisting of the estimated regression coefficients,
#the estimated standard errors of the regression coefficients,
#the residuals, the fitted values, 'dep', 'indep', and 'y' which is 'dep'
#but as matrix.

*The Inputs

The function has 4 inputs specified. 'data', 'dep', 'indep', and 'intercept'.
'data' is the dataset you wish to use to construct your model, 'dep' is the 
name of the vector corresonding to the dependent variable, 'indep' is the vector
of names for the vectors corresponding to the independent variables,
'intercept' determines whether the model has an intercept or not.

*If statements

To make sure that the input of the function is suitable some if-statements
have been constructed after the specification of the inputs. The first 
if-statement controls that the input for 'data' is a tibble. The second 
if-statement controls that the input 'dep' consist of a character vector with 
names in the tibble of 'data'. The third if-statemnet controls that there is 
only one name in the character of 'dep'. The fourth if-statement controls that 
there is at least on element in character vector of names in the input of 'dep'. 
The fifth if-statement controls that the entries in the character of 'indep' are
all in the names of 'data'.

*Transforming 'dep' and 'indep' into matrices 'y' and 'X'.

*if-statements adding vector of 1s if intercept is included.

*Algebra

*Output

*Print function

*Plot function

```{r}
linear_regression <- function(data, dep, indep, intercept = TRUE)
{
  
  if (!is_tibble(data)) {stop("the input 'data' must be a tibble")}
  
  if (!dep %in% names(data)) {stop("No dependent variable with that name in your tibble of data")} #If there is no variable named 'dep' in 'data' stop.
  
  if(!(is.character(dep) && length(dep) == 1)) {
    stop("'dep' must be a character vector of length 1")}
  
  if (!(is.character(indep) && length(indep) > 0)) {
    stop("'indep' must be a character vector of length 1 or greater.")}
  
  if (!all(indep %in% names(data))) {stop("No independet variables with that name in your tibble of data")}
  
  X <- as.matrix(data[,indep])
  y <- as.matrix(data[[dep]])
  
  if (intercept == TRUE) {
    rows_in_vector_of_1s <- nrow(X)  # Get the number of rows for the vector of 1s
    Intercept <- rep(1, rows_in_vector_of_1s) #Creating a vector of 1s
    X <- cbind(Intercept, X) #Input matrix when intercept is used
  }
  
  if (!is.logical(intercept)) {stop("The input 'intercept' must be a logical variable")}
  
  
  XT <- t(X) #Transpose of the matrix of independent variables
  XTXI <- solve(XT %*% X) #Inverse of the X^T times X.
  beta <- XTXI %*% XT %*% y #Calculation of beta
  fitted <- X%*%beta #Calculation of fitted values
  residuals <- y-fitted #Calculation of residuals
  sumofres <- sum(residuals^2) #sum of the square residuals
  sigma <- sqrt(sumofres/(nrow(X)-ncol(X))) #sigma is the squareroot of the sum of the square residuals divided by of number of rows of the matrix of indepdendent variables minus its rows
  vector_of_variances_of_betas <- diag((sigma^2)*XTXI) #Calculation of vector of variances
  se <- sqrt(vector_of_variances_of_betas) #We squareroot it to get the standard errors of the betas.
  
  output <- list(beta = beta, se = se, residuals = residuals, fitted = fitted, sigma = sigma, dep = dep, indep = indep, y = y)
  return(output)} #Our linear regression function

print <- function(the_list){ #function for printing
  beta <- the_list$beta
  se <- the_list$se
  my_matrix <- rbind(t(beta), se)
  rownames(my_matrix) <- c("coefficient", "se")
  return(my_matrix)}

plot <- function(the_list){ #function for plot
  fitted <- the_list$fitted
  y <- the_list$y
  newdf <- data.frame(Actual = y, Fitted = fitted)
  ggplot(newdf, aes(x = (1:length(y)))) +
    geom_point(aes(y = Actual, color = "Actual"), size = 1) +
    geom_point(aes(y = Fitted, colour = "Fitted"), size = 1) +
    labs(title = "Actual and fitted values (sorted)", x = "Order", y = "Value") +
    scale_color_manual(values = c("Actual" = "turquoise", "Fitted" = "orange", name = "Variable")) +
    theme_minimal()
}



obj <- linear_regression(ggplot2::economics,
                         dep = "pce",
                         indep = c("unemploy", "pop", "psavert"))



print(obj)
plot(obj)

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
